### 2019-06-26 manuscript.Rmd
move files

In our earlier work [@fphar201800588], we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from @jansen1999analysis and @owen2013better. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09} and related standard \proglang{C} compiler to solve the ODE function. In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE. The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}. The \proglang{C} compiler had originated in Linux or MacOS. The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as,

```{r, eval=F}
Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
Sys.setenv(PATH = paste(Rtools.path, Sys.getenv("PATH"), sep=";"))
```

@@

We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative]. Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity]. The eFAST method compute the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis. This type of GSA approach and SI also know as Sobol method and Sobol index. Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement. This can be written as:

@@

## Pharmacokinetic modeling

@@

The uncertainty analysis is a crucial modeling process with sensitivity analysis. Sometimes, it's difficult to have informative information to set up the parameter distribution. The wider range of parameter distribution might cause the computation error at the extreme value. The narrower range might cause the incorrect calibrated result of the model parameter. To prevent this problem. The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process. Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.

@@

The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model. The model mainly includes two state variables that are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}). The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance. The \code{AUC} is the calculated area under the curve of the rest body ($mol \cdot h/L$).

>>>>

In our earlier work [@fphar201800588], we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from @jansen1999analysis and @owen2013better. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function. In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE. The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}. The \proglang{C} compiler had originated in Linux or MacOS. The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as,

```{r, eval=T}
Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
Sys.setenv(PATH = paste(Rtools.path, Sys.getenv("PATH"), sep=";"))
```

@@

We adopted eFAST, a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative]. Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity]. The eFAST method compute the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis. This type of GSA approach and SI also know as Sobol method and Sobol index. Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement. This can be written as:

@@

## Modeling

@@

The uncertainty analysis is a crucial modeling process with sensitivity analysis. Sometimes, it's difficult to have informative data to set up the parameter distribution. The wider range of parameter distribution might cause the computation error at the extreme value. The narrower range might cause the incorrect calibrated result of the model parameter. To prevent this problem. The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process. Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output. \pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result. The general workflow of uncertainty and sensitivityanalysis is illustrated in Figure \ref{fig:Workflow}.

```{r workflow, echo = FALSE, out.width = "100%", fig.cap = '\\label{fig:Workflow}Workflow for performing uncertainty and sensitivity analysis in \\pkg{pksensi}.'}
knitr::include_graphics("diagram/workflow.png")
```

@@

The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model. The model mainly includes two state variables that are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}). The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance. The \code{AUC} is the calculated area under the curve of the rest body ($mol \cdot h/L$).
